Structural Stability of Random Dynamical Systems

### 2014-09-04 journal.tex
clarify 2x2 case and add figures

Suppose that we resample $b$.  Since the trace does not involve $b$,
that condition will be satisfied automatically and we only need to
examine the determinant.  We will consider seperately the cases of $d
< 0$ and $d > 0$.  When $d > 0$, the condition becomes $ac/d > b$,
which is satisfiable when $ac > -d$ and results in probability $(ac/d
+ 1)/2$ of remaining stable.  When $d < 0$, the condition becomes
$ac/d < b$, which is satisfiable when $ac > d$, so we have probability
$(1 - ac/d)/2$.

@@

In order to evaluate the integrals, we begin with the
integrals over $b$, which are particularly easy because none of the
integrands depends upon $b$.  Once we do that and make the change of
variables $d \mapsto -d$ in the second integrals in the numerator and
denominator, we can combine terms and simplify the expression to the
following:
$$
{\int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2 + \int_0^1 dd \int_{d > ac} da\,dc\, 4 \over
{2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)} + 2 \int_0^1 dd \int_{d > ac} da\,dc\, 2}
$$
For $d > 0$ we have
$$
\int_{-d<ac<d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c) 
%\over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
$$
and 
$$
\int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$

@@

Suppose that we resample $a$.  Then we have to satisfy the conditions
$a + c < 0$ and $ac > bd$ for fixed values of $b,c,d$.  We will
consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
inequalities become $a < -c$ and $a > bd/c$.  For these to be
consistent, we must have $-c > bd/c$, or $bd < -c^2$ and, when we do,
we have a probability $(-c-bd/c)/2$ of reasmpling an $a$ that keeps
our matrix stable.  When $c < 0$, the conditions become $a < -c$ and
$a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
< -c$.  Then the probability of resampling an $a$ which keeps the
matrix stable is $(\min(-c, bd/c) + 1)/2$.

@@

>>>>

Suppose that we resample $b$.  Since the trace does not involve $b$,
that condition will be satisfied automatically and we only need to
examine the determinant.  Thus, we have the inequalities $ac > bd$ and $-1 < b < 1$.  We will consider seperately the cases of $d
< 0$ and $d > 0$.  When $d > 0$, the first condition becomes $ac/d > b$.
Since $b > -1$, this can only be satisfied when $ac/d > -1$, i.e. when $ac > -d$.  In the subcase that $ac < d$, we then have the limits $-1 < b < ac/d$ and in the subcase $ac < d$, we have the limits $-1 < b < 1$.  When $d < 0$, the condition becomes
$ac/d < b$, which is satisfiable when $ac > d$.  In the subcase where $ac > -d$, we have $ac/d < b < 1$ and, in the subcase where $ac < -d$, we have $-1 < b <
1$.

@@

In order to evaluate the integrals, we begin with the
integrals over $b$, which are particularly easy because none of the
integrands depends upon $b$.  Once we do that and make the change of
variables $d \mapsto -d$ in the second integrals in the numerator and
denominator, we can combine terms and simplify the expression to the
following:
$$
\frac{\int_0^1 dd \int_{ac > d} da\,dc\, 4 + \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)^2}{2 \int_0^1 dd \int_{ac > d} da\,dc\, 2 + 2 \int_0^1 dd \int_{d > ac > -d} da\,dc ({ac \over d} + 1)}
$$

\begin{figure}[!ht]
\centering
\noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-b.pdf}
\caption{{\bf Resampling b.} (A) The region defined by $ac > d$ (B) The region defined by $d > ac > -d$}
\label{fig:2x2-resample-b}
\end{figure}

For $ac > d$ we divide the $a$-axis into two parts indicated in Fig. \ref{fig:2x2-resample-b}A
$$
\int_{ac>d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{-1}^{d \over a} dc \, f(a,c) + \int_{d}^{1} da \int_{d \over a}^1 dc \, f(a,c)$$
for the numerator $f(a,c) = 4$ and the integral is equal to 2. For $d > ac > -d$ we divide the $a$-axis into three parts indicated in Fig. \ref{fig:2x2-resample-b}B
$$
\int_{d > ac > -d} da\,dc \, f(a,c) = \int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc \, f(a,c) + \int_{-d}^{d} da \int_{-1}^{1} dc \, f(a,c) + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc \, f(a,c)
% \over {\int_{-1}^{-d} da \int_{d \over a}^{- d \over a} dc + \int {-d}^{d} da \int_{-1}^{1} dc + \int_{d}^{1} da \int_{-d \over a}^{d \over a} dc}
$$
for the numerator $f(a,c) = (\frac{ac}{d} + 1)^2$ and the integral is equal to $\frac{32}{9}$ and for the denominator $f(a,c) = \frac{ac}{d} + 1$ and the integral is equal to $3$. Therefore
$$P\left(\begin{pmatrix}
a & b' \\
d & c
\end{pmatrix} \textrm{stable } \bigg| \begin{pmatrix}
a & b \\
d & c
\end{pmatrix} \textrm{stable } \right) = \frac{25}{36} \approx 0.69.$$

@@

Now suppose we resample $a$.  Then we have to satisfy the conditions
$a + c < 0$, $ac > bd$, and $-1 < a < 1$ for fixed values of $b,c,d$.  We will
consider seperately the cases $c > 0$ and $c < 0$.  When $c > 0$, our
inequalities become $a < -c$ and $a > bd/c$.  For these to be
consistent, we must have $-c > bd/c$, or $bd < -c^2$.  In the subcase where $bd < -c$, we have $-1 < a < -c$ and, in the subcase where $-c < bd < -c^2$, we have $bd/c < a < -c$. When $c < 0$, the conditions become $a < -c$ and
$a < bd/c$, which will have a solution as long as $bd/c > -1$, or $bd
< -c$.  In the subcase where $bd < -c^2$, we have $-1 < a < bd/c$ and, in the subcase where $bd > -c^2$, we have $-1 < a < -c$.

@@

\begin{figure}[!ht]
\centering
\noindent\includegraphics[width=0.5\columnwidth]{fig/2x2-resample-a.pdf}
\caption{{\bf Resampling a.} (A) The region defined by $bd < -c$ (B) The region defined by $-c < bd < -c^2$ (C) The region defined by $bd < -c^2$ (D) The region defined by $bd > -c^2$}
\label{fig:2x2-resample-a}
\end{figure}
